Depression and suicidality as costly signals of need

### 2021-03-29 Paper.rmd
Added doi for 2020 data
Added text to the methods section

---
title: "Depression and suicidality as costly signals of need"
author: "Michael Gaffney, Kai Adams, Kristen L. Syme, and Edward H. Hagen"
date: '`r format(Sys.Date(), "%B %d, %Y")`'
always_allow_html: true
output: 
  bookdown::html_document2:
    number_sections: false
bibliography: MTBib.bib
csl: apa-6th-edition.csl
---

@@

## notes: 
add funding info somewhere
update MTurk (Ed email links + mine)
Find a place for the predictions and preregistration 
go back and emohasize conflit and PI in the methods (at least in the pilot studies portion)

@@

@@

@@

@@

@@

The total effect of the *Depression* signal (treatment) vs. *Verbal request* (control) on the likelihood of *Action* was to increase it by `r 100*signif(med1b$tau.coef, 3)` points from T1 to T2 (on the original 100-point scale). Of this increase, `r 100*signif(med1b$n.avg, 2)`% was mediated by the increased *Belief* that the victim was telling the truth. See Figure \@ref(fig:mediation).

@@

@@

Briefly, Participants read a vignette that put them in situations in which an imagined sister was requesting a loan of $50,000 from the participant that was originally saved for the participant's imagined daughter’s college education. The sister claims the money is needed for medical treatment for her own child, but she has incentives to deceive. The manipulations were signals very similar to those used in this study, and the outcomes were belief that the sister really needed the money and intention to load her the money. See XX for the full write-up. See https://osf.io/3rg8b for preregistration of the pilot study.

>>>>

---
title: "Depression and suicidality as costly signals of need"
author: "Michael R. Gaffney, Kai Adams, Kristen L. Syme, and Edward H. Hagen"
date: '`r format(Sys.Date(), "%B %d, %Y")`'
always_allow_html: true
output: 
  bookdown::html_document2:
    number_sections: false
bibliography: MTBib.bib
csl: apa-6th-edition.csl
---

@@

## notes: 
add funding info somewhere
update MTurk (Ed email links + mine)
Find a place for the predictions and preregistration 
go back and emohasize conflit and PI in the methods (at least in the pilot studies portion)
add abstract
add conclusion

@@

## Data availibility

http://doi.org/10.5281/zenodo.4637904

@@

1,950 participants started the study by clicking the link to Qualtrics (1,213 American and 737 Indian). After removing participants who did not finish the survey, failed attention checks, or moved through the study at an unrealistc pace, our sample dropped to N = 1,240 (937 American and 303 Indian). Basic summary statistics on the remaining N = 1,240 participants (Male = 759, Female = 479) can be found in table XX.

@@

Power analyses prior to data collection showed that high proportions of participants scoring extremely high on belief or likelihood of helping at T1 would result in low power for this study's design, a problem which we appeared to largely avoid (see figure XX).

Although not a part of our pre-registered predictions, both T1Belief and T1Action were highly correlated across the four vignettes (enter correlation coefficients), consistent with help being worth providing if the signaler's claims were true (see figure XX).

@@

As a whole, our predictions for the effect of signal cost on belief and likelihood of helping T2 were met, with costlier signals eliciting greater pro-signaler responses. There was also support for an interaction between signal and T1Belief or T1Action, although the effect was not large (See figure 2). However, the effect of signal cost varied by vignette, with signal type having the greatest effect in the "basketball coach" and "brother-in-law" vignettes. When ordered according to cost, the greatest jump in support tended to occur between crying and mild-depression, with the next largest jump tending to occur between depression and suicide attempt. Overall, there was little difference between the effect of mild depression and the effect of depression.

@@

The total effect of the *Depression* signal (treatment) vs. *Verbal request* (control) on the likelihood of *Action* was to increase it by `r 100*signif(med1b$tau.coef, 3)` points from T1 to T2 (on the original 100-point scale). Of this increase, `r 100*signif(med1b$n.avg, 2)`% was mediated by the increased *Belief* that the victim was telling the truth. See Figure \@ref(fig:mediation). The total effect of the *Suicide attempt* signal (treatment) vs. *Verbal request* (control) on the likelihood of *Action* was to increase it by `r 100*signif(med4b$tau.coef, 3)` points from T1 to T2. Of this increase, `r 100*signif(med4b$n.avg, 2)`% was mediated by the increased *Belief* that the victim was telling the truth (Check if med4 or med4b is correct; they look identical to me).

@@

We also examined the effect of signaling on how participants perceived the signaler's emotional state. To do this, we created four new variables: LowMood (T1 and T2) and Manipulative (T1 and T2). This was done by adding the dummy variables for whether a participant thought the signaler was depressed, distressed or sad for the LowMood variables and by adding the dummy variables for whether they thought the particpant was Devious or Jealous for the Manipulative variables. In all instances, 0 indicated a participant did not select the emotion, while 1 indicated selection.

We then plotted the mean change seen in the data in addition to the change seen in 500 bootstrap resamples (see figure 6).

@@

Briefly, Participants read a vignette that put them in situations in which an imagined sister was requesting a loan of $50,000 from the participant that was originally saved for the participant's imagined daughter’s college education. The sister claims the money is needed for medical treatment for her own child, but she has incentives to deceive. The manipulations were signals very similar to those used in this study, and the outcomes were belief that the sister really needed the money and intention to load her the money. See XX for the full write-up. See https://osf.io/3rg8b for preregistration of the pilot study. See http://doi.org/10.5281/zenodo.4637883 for pilot data.
